---
title: "R Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
install.packages("clusterGeneration")
install.packages("mvtnorm")
install.packages("rmutil")
# package for checking whether the matrix is positive definite
install.packages("matrixcalc")
library(clusterGeneration)
library(MASS)
library(mvtnorm)
library(Matrix)
library(matrixcalc)
```

f: a function defined to return the f exceedances.
```{r}
# x: a vector containing the extreme variables that are condition on
# gamma: a vector of parameters attached with the extreme variables(X4 X5) that we condition on
# delta: control parameter here we will fix to 2
f <- function(x, gamma, delta) (sum(gamma*((gamma*x)^delta)))^(1/delta)
```

pro.lik: a function defined to perform profile likelihood estimation given more than one extremes.
```{r}
# theta: a vector of parameters of interest including gammas, beta, c
# dat: a data matrix containing the extremes(X4 X5) and other variable(X6) 
# p: the column number of extreme
# thresh: a selected threshold 

pro.lik <- function(theta, dat, p, u){
  
  # extract the index of exceedances
  exc <- apply(dat[,-p], 1, function(x){
    if(sum(x > u) > 0){
      return(1)}
    else{
      return(0)}
  })
  
  # extract the rows of exceedances
  dat.exc <- dat[exc == 1,]
  n <- nrow(dat.exc)
  # extract the response variable: not the extremeX6
  resp <- dat.exc[,p]
  # extract the explanatory variables
  var <- dat.exc[,-p]
  d <- ncol(var)
  
  # apply additive log-ratio transformation
  theta[1:d-1] <- log(theta[1:d-1]/theta[d])
  # add a zero to get a d-dimensional vector
  theta[1:d] <- append(theta[1:d-1], 0) 
  # inverse to obtain gammas that sum up to 1
  gamma <- exp(theta[1:d])/sum(exp(theta[1:d]))
  
  # specify the parameter beta and c
  beta <- exp(theta[d+1])/(1+exp(theta[d+1]))
  c <- exp(theta[d+2])

  # call the function f to get the values of f.exc
  f.exc <- apply(var, 1, f, gamma=gamma, delta=2)
  # derive the expressions of estimated mu and sigma
  resid <- (resp - c*f.exc)/(f.exc)^beta
  mu <- mean(resid)
  sig <- sd(resid)
  
  # compute the profile log-likelihood
  #pl <- -n*log(sig)-beta*sum(log(f.exc))
  pl <- -(n/2)*log(sig) - sum(log(f.exc))
  
  return(pl)
}
```

output: a function to obtain the outputs of maximizing pro.lik
```{r}
output <- function(theta, dat, p, u){
  
  # extract the index of exceedances
  exc <- apply(dat[,-p], 1, function(x){
    if(sum(x > u) > 0){
      return(1)}
    else{
      return(0)}
  })
  
  # extract the rows of exceedances
  dat.exc <- dat[exc == 1,]
  n <- nrow(dat.exc)
  # extract the response variable: not the extremeX6
  resp <- dat.exc[,p]
  # extract the explanatory variables
  var <- dat.exc[,-p]
  d <- ncol(var)
  
  theta[1:d-1] <- log(theta[1:d-1]/theta[d])
  # add a zero to get a d-dimensional vector
  theta[1:d] <- append(theta[1:d-1], 0) 
  # plug in the estimated theta to get estimators
  gamma <- exp(theta[1:d])/sum(exp(theta[1:d]))
  beta <- exp(theta[d+1])/(1+exp(theta[d+1]))
  c <- exp(theta[d+2])
  
  # call the function f to get the values of f.exc
  f.exc <- apply(var, 1, f, gamma=gamma, delta=2)
  # derive the expressions of estimated mu and sigma
  resid <- (resp - c*f.exc)/(f.exc)^beta
  mu <- mean(resid)
  sig <- sd(resid)
  
  output <- list()
  output$gamma <- gamma
  output$beta <- beta
  output$c <- c
  output$resid  <- resid
  output$f.exc <- f.exc
  output$resp <- resp

  return(output)
}
```

pllik_Lk: a function to perform profile likelihood estimation given only one extreme.
```{r}
pllik_Lk <- function (x, dat, which, u) {
    d <- ncol(dat)
    n <- nrow(dat[dat[,which]>u,])    
    
    alphak            <- (exp(x[seq(1,(d-1),by=1)])/(1+exp(x[seq(1,(d-1),by=1)])))
    betak             <- exp(x[seq(d,2*(d-1),by=1)])/(1+exp(x[seq(d,2*(d-1),by=1)]))
    dat_E <- lapply(1:(d-1),
                    function(J){
                        K       <- ((1:d)[-which])[J]
                        Y       <- dat[dat[,which]>u,K]
                        X       <- dat[dat[,which]>u,which]
                        Z       <- (Y-alphak[J]*X)/X^(betak[J])
                        sumX    <- sum(betak[J]*log(X))
                        sig     <- var(Z)
                        Zb      <- list()
                        Zb$Z    <- Z
                        Zb$sig  <- sig
                        Zb$sumX <- sumX
                        return(Zb)
                     })
    Z       <- do.call("cbind",lapply(1:(d-1), function(x) dat_E[[x]]$Z))
    sumX    <- sum(unlist(lapply(1:(d-1), function(x) dat_E[[x]]$sumX)))
    sumX    <- unlist(lapply(1:(d-1), function(x) dat_E[[x]]$sumX))
    sig     <- unlist(lapply(1:(d-1), function(x) dat_E[[x]]$sig))
    muZ     <- colMeans(Z)    
    Zminmu  <- sweep(Z, 2, muZ, FUN="-")
    Sigma   <- (t(Zminmu)%*%Zminmu)/n
    Pl      <- -(n/2)*log(det(Sigma)) - sum(sumX)
    
    return(Pl)
}
```

HTfit.model.output.dev: a function to obtain the outputs of maximizing pllik_Lk.
```{r}
HTfit.model.output.dev <- function(dat, u, init){
    d   <- ncol(dat)
    nE  <- unlist(lapply(as.list(1:d), function(x)  sum(dat[,x]>u)))
    A   <<- Matrix(1, nrow=d, ncol=d)
    B   <<- Matrix(0, nrow=d, ncol=d)
    out <- lapply(as.list(1:d), function(k) {
        X    <- (dat[dat[,k]>u, k])
        opt  <- optim(par=init, pllik_Lk,
                      dat=dat,u=u, which=k,
                      control=list(fnscale=-1,maxit=1e5))
        alphak   <- (exp(opt$par[1:(d-1)])/(1+exp(opt$par[1:(d-1)])))
        A[k,-k ] <<- alphak
        betak    <- exp(opt$par[d:(2*(d-1))])/(1+exp(opt$par[d:(2*(d-1))]))
        B[k,-k ] <<- betak
        Z <- Matrix(0, nrow=nE[k], ncol=d)
        Z[,-k] <- do.call("cbind", lapply(as.list(1:(d-1)), function(y) {
            Y <- as.matrix((dat[dat[,k]>u, -k]))[,y]
            resid <- (Y - alphak[y]*X)/X^(betak[y])
            return (resid)
        }))
        output <- list()
        output$opt    <- opt
        output$Z      <- Z
        output$alphak <- alphak
        output$betak  <- betak
        return(output)
    }
    )
    model_output <- list()
    model_output$A <- A
    model_output$B <- B
    model_output$Z <- lapply(out, function(x) x$Z)
    
    return(model_output)
}
```

