---
title: "2-tail graph"
output: html_document
---

data: a function to generate samples from Gaussian multivariate distribution
datapit: a function to perform PIT to transform the data to exp(1) margins
```{r}
data <- function(N, d) {
  mu <- rep(0, d)
  Sigma <- Sigma
  X_Gaussian <- mvrnorm(n = N, mu = mu, Sigma = Sigma) 
  return(X_Gaussian)
}

datapit <-function(dat) {
  edf <- apply(dat, 2, function(x) rank(x)/(1+length(x))) 
  X <- apply(edf, 2, function(x) qexp(x,1)) 
  return(X)
}
```

Tail graph : generate a random positive definite correlation matrix based on the graphical structure.
```{r}
set.seed(2020)
# generate a random positive definite correlation matrix
Sigma <- rcorrmatrix(6)
# specify the graphical structure of the tail graph
for (j in c(3:6) ){
  Sigma[1,j] <- 0
}
for (i in c(1:3) ){
  Sigma[i,6] <- 0
}
# force the matrix to be symmetric
Sigma[lower.tri(Sigma)] <- t(Sigma)[lower.tri(Sigma)]

# check whether the the matrix is positive definite
is.positive.definite(Sigma)
```

Apply Monte Carlo method and create lists to store the generated data.
```{r}
# Monte Carlo iteration: M=500
M <- 500
# specify the sample size
N <- 1000
# specify the dimensions
d <- 6

# create lists to store the generated data for Monte Carlo iterations
X_G <- vector('list', M)
X_PIT <- vector('list', M)
U <- vector('list', M)
# call the funtion to perform data generation
for(I in 1:M){
  X_G[[I]] <- data(N, d)
  X_PIT[[I]] <- datapit(X_G[[I]])
  U[[I]] <- quantile(X_PIT[[I]], 0.85)  
}
```

Create lists to store the model outputs.
```{r}
# for initial clique
IC <- vector('list', M)
IC_fit <- vector('list', M)
A_IC <- vector('list', M)
B_IC <- vector('list', M)
Z_IC <- vector('list', M)

# for the second clique
SC <- vector('list', M)
SC_fit <- vector('list', M)
A_SC <- vector('list', M)
B_SC <- vector('list', M)
Z_SC <- vector('list', M)

# for the third clique
TC <- vector('list', M)
TC_fit <- vector('list', M)
TC_OUTPUT <- vector('list', M)
gamma_TC <- vector('list', M)
A_TC <- vector('list', M)
B_TC <- vector('list', M)
Z_TC <- vector('list', M)
EXC_TC <- vector('list', M)
R_TC <- vector('list', M)
```

Fit the initial clique: X245 | X3
```{r}
for(I in 1:M){
  IC[[I]] <- X_PIT[[I]][,c(2:5)]
  IC_fit[[I]] <- HTfit.model.output.dev(IC[[I]], U[[I]], init=rep(c(0.5,0.1),4))
  A_IC[[I]] <- IC_fit[[I]]$A
  B_IC[[I]] <- IC_fit[[I]]$B
  Z_IC[[I]] <- IC_fit[[I]]$Z
}
# obtain Monte Carlo estimators
MA_IC <- Reduce("+", A_IC) / length(A_IC)
MA_IC <- as.matrix(MA_IC)
MB_IC <- Reduce("+", B_IC) / length(B_IC)
MB_IC <- as.matrix(MB_IC)
```

```{r}
# obtain Monte Carlo estimators for Z
z2_list <- NULL
z3_list <- NULL
z4_list <- NULL
z5_list <- NULL
i <- 1
for (I in 1:M) {
    z2_list[[i]] <- as.matrix(Z_IC[[I]][[1]])
    z3_list[[i]] <- as.matrix(Z_IC[[I]][[2]])
    z4_list[[i]] <- as.matrix(Z_IC[[I]][[3]])
    z5_list[[i]] <- as.matrix(Z_IC[[I]][[4]])
    i <- i + 1
}
Z_list <- c(list(z2_list),list(z3_list), list(z4_list), list(z5_list))
MZ2345 <- NULL
for (l in 1:4) {
  MZ2345[[l]] <- Reduce("+", Z_list[[l]]) / length(Z_list[[l]])
  MZ2345[[l]] <- as.matrix(MZ2345[[l]])
}
```

Fit the second clique: X1 | X2
```{r}
for(I in 1:M){
  SC[[I]] <- X_PIT[[I]][,c(1:2)]
  SC_fit[[I]] <- HTfit.model.output.dev(SC[[I]], U[[I]], init=rep(c(0.5,0.1),2))
  A_SC[[I]] <- SC_fit[[I]]$A
  B_SC[[I]] <- SC_fit[[I]]$B
  Z_SC[[I]] <- SC_fit[[I]]$Z
}

# obtain Monte Carlo estimators
MA_SC <- Reduce("+", A_SC) / length(A_SC)
MA_SC <- as.matrix(MA_SC)
MB_SC <- Reduce("+", B_SC) / length(B_SC)
MB_SC <- as.matrix(MB_SC)
```

```{r}
# obtain Monte Carlo estimators for Z
z1_list <- NULL
z22_list <- NULL
i <- 1
for (I in 1:M) {
    z1_list[[i]] <- as.matrix(Z_SC[[I]][[1]])
    z22_list[[i]] <- as.matrix(Z_SC[[I]][[2]])
    i <- i + 1
}
Z2_list <- c(list(z1_list),list(z22_list))
MZ12 <- NULL
for (l in 1:2) {
  MZ12[[l]] <- Reduce("+", Z2_list[[l]]) / length(Z2_list[[l]])
  MZ12[[l]] <- as.matrix(MZ12[[l]])
}
```

Fit the third clique: X6 | X45
```{r}
for(I in 1:M){
  TC[[I]] <- X_PIT[[I]][,c(4:6)]
  TC_fit[[I]] <- optim(par=c(0.1, 0.1, 0, 0), fn=pro.lik, dat=TC[[I]], p=3, u=U[[I]], 
                       method="Nelder-Mead", control=list(fnscale=-1, maxit=5000))
  TC_OUTPUT[[I]] <- output(theta=TC_fit[[I]]$par, dat=TC[[I]], p=3, u=U[[1]]) 
  gamma_TC[[I]] <- TC_OUTPUT[[I]]$gamma
  B_TC[[I]] <- TC_OUTPUT[[I]]$beta
  A_TC[[I]] <- TC_OUTPUT[[I]]$c
  Z_TC[[I]] <- TC_OUTPUT[[I]]$resid
  EXC_TC[[I]] <- TC_OUTPUT[[I]]$f.exc
  R_TC[[I]] <- TC_OUTPUT[[I]]$resp
}
# obtain Monte Carlo estimators
MA_TC <- Reduce("+", A_TC) / length(A_TC)
MA_TC <- as.matrix(MA_TC)
MB_TC <- Reduce("+", B_TC) / length(B_TC)
MB_TC <- as.matrix(MB_TC)
Mgamma_TC <- Reduce("+", gamma_TC) / length(gamma_TC)
Mgamma_TC <- Mgamma_TC
MZ_TC <- Reduce("+", Z_TC) / length(Z_TC)
MZ_TC <- as.matrix(MZ_TC)
```

Fit the third clique to obtain the parameters for X45 | X6:
```{r}
C <- vector('list', M)
C_fit <- vector('list', M)
A_C <- vector('list', M)
B_C <- vector('list', M)
Z_C <- vector('list', M)

for(I in 1:M){
  C[[I]] <- X_PIT[[I]][,c(4:6)]
  C_fit[[I]] <- HTfit.model.output.dev(C[[I]], U[[I]], init=rep(c(0.5,0.1),3))
  A_C[[I]] <- C_fit[[I]]$A
  B_C[[I]] <- C_fit[[I]]$B
  Z_C[[I]] <- C_fit[[I]]$Z
}

# obtain Monte Carlo estimators
MA_C <- Reduce("+", A_C) / length(A_C)
MA_C <- as.matrix(MA_C)
MB_C <- Reduce("+", B_C) / length(B_C)
MB_C <- as.matrix(MB_C)

z44_list <- NULL
z55_list <- NULL
z66_list <- NULL
i <- 1
for (I in 1:M) {
    z44_list[[i]] <- as.matrix(Z_C[[I]][[1]])
    z55_list[[i]] <- as.matrix(Z_C[[I]][[2]])
    z66_list[[i]] <- as.matrix(Z_C[[I]][[3]])
    i <- i + 1
}
Z456_list <- c(list(z44_list),list(z55_list), list(z66_list))
MZ456 <- NULL
for (l in 1:3) {
  MZ456[[l]] <- Reduce("+", Z456_list[[l]]) / length(Z456_list[[l]])
  MZ456[[l]] <- as.matrix(MZ456[[l]])
}
```

Construct the full beta matrix:
```{r}
B <- matrix(0, 6, 6)
B[1:nrow(MB_SC), 1:ncol(MB_SC)] <- MB_SC
B[2:(nrow(MB_IC)+1), 2:(ncol(MB_IC)+1)] <- MB_IC
B[1,3:6]<- MB_TC
B[2,6]<- MB_TC
B[3,1]<- MB_TC
B[3,6]<- MB_TC
B[4,1]<- MB_TC
B[4,6]<- MB_TC
B[5,1]<- MB_TC
B[5,6]<- MB_TC
B[6,1:5]<- MB_TC
B
```

Construct the full alpha matrix:
```{r}
A <- matrix(0, 6, 6)
A[1:nrow(MA_SC), 1:ncol(MA_SC)] <- MA_SC
A[2:(nrow(MA_IC)+1), 2:(ncol(MA_IC)+1)] <- MA_IC
A[6, 4:6] <- MA_C[3,]
A[4:6, 6] <- MA_C[,3]

X2 <- NULL
X24 <-NULL
X25 <-NULL
Z24 <- MZ2345[[2]][,3]
Z25 <- MZ2345[[2]][,4]
i <- 1
for (I in 1:M) {
  X2[[i]] <- IC[[I]][,1][IC[[I]][,1]>U]
  X24[[i]] <- A[2,4]*X2[[i]]+(X2[[i]]^B[2,4])*Z24
  X25[[i]] <- A[2,5]*X2[[i]]+(X2[[i]]^B[2,5])*Z25
  i <- i + 1
}

X3 <- NULL
X34 <-NULL
X35 <-NULL
Z34 <- MZ2345[[3]][,3]
Z35 <- MZ2345[[3]][,4]
i <- 1
for (I in 1:M) {
  X3[[i]] <- IC[[I]][,2][IC[[I]][,2]>U]
  X34[[i]] <- A[3,4]*X3[[i]]+(X3[[i]]^B[3,4])*Z34
  X35[[i]] <- A[3,5]*X3[[i]]+(X3[[i]]^B[3,5])*Z35
  i <- i + 1
}

aplha26 <- NULL
aplha36 <- NULL
i <- 1
for (I in 1:M) {
  alpha26[[i]] <- mean(apply(cbind(X24[[I]], X25[[I]]), 1, f, gamma=gamma_TC[[I]], delta=2))
  alpha36[[i]] <- mean(apply(cbind(X34[[I]], X35[[I]]), 1, f, gamma=gamma_TC[[I]], delta=2))
  i <- i+1
}

A[2,6] <- MA_TC*(mean(alpha26))
A[3,6] <- MA_TC*(mean(alpha36))
A[6,2] <- A[6,4]*A[6,5]*A[4,2]*A[5,2]
A[6,3] <- A[6,4]*A[6,5]*A[4,3]*A[5,3]

A[3,1] <- A[3,2]*A[2,1]
A[4,1] <- A[4,2]*A[2,1]
A[5,1] <- A[5,2]*A[2,1]
A[6,1] <- A[6,2]*A[2,1]

A[1,3] <- A[1,2]*A[2,3]
A[1,4] <- A[1,2]*A[2,4]
A[1,5] <- A[1,2]*A[2,5]
A[1,6] <- A[1,2]*A[2,6]

A
```



