---
title: "4-block graph"
output: html_document
---

Block graph : generate a random positive definite correlation matrix based on the graphical structure.
```{r}
set.seed(2020)
# Generate a random correlation matrix based on random partial correlations
Sigma <- rcorrmatrix(5)
# Specify the graphical structure
for (j in c(4:5) ){
  Sigma[1,j] <- 0
}
for (i in c(4:5) ){
  Sigma[2,i] <- 0
}
# Force the matrix to be symmetric
Sigma[lower.tri(Sigma)] <- t(Sigma)[lower.tri(Sigma)]
# check whether the the matrix is positive definite
is.positive.definite(Sigma)
```

Apply Monte Carlo method and create lists to store the generated data.
```{r}
# Monte carlo iteration: M=500
M <- 500
N <- 1000
d <- 5
X_G <- vector('list', M)
X_PIT <- vector('list', M)
U <- vector('list', M)

for(I in 1:M){
  X_G[[I]] <- data(N, d)
  X_PIT[[I]] <- datapit(X_G[[I]])
  U[[I]] <- quantile(X_PIT[[I]], 0.85)  
}
```

Create lists to store the model outputs.
```{r}
# for initial clique
IC <- vector('list', M)
IC_fit <- vector('list', M)
A_IC <- vector('list', M)
B_IC <- vector('list', M)
Z_IC <- vector('list', M)

# for the second clique
SC <- vector('list', M)
SC_fit <- vector('list', M)
A_SC <- vector('list', M)
B_SC <- vector('list', M)
Z_SC <- vector('list', M)
```

Fit the initial clique: X12 | X3
```{r}
for(I in 1:M){
  IC[[I]] <- X_PIT[[I]][,c(1:3)]
  IC_fit[[I]] <- HTfit.model.output.dev(IC[[I]], U[[I]], init=rep(c(0.5,0.1),3))
  A_IC[[I]] <- IC_fit[[I]]$A
  B_IC[[I]] <- IC_fit[[I]]$B
  Z_IC[[I]] <- IC_fit[[I]]$Z
}
# obtain Monte Carlo estimators
MA_IC <- Reduce("+", A_IC) / length(A_IC)
MA_IC <- as.matrix(MA_IC)
MB_IC <- Reduce("+", B_IC) / length(B_IC)
MB_IC <- as.matrix(MB_IC)
```

```{r}
# obtain Monte Carlo estimators for Z
z1_list <- NULL
z2_list <- NULL
z3_list <- NULL

i <- 1
for (I in 1:M) {
    z1_list[[i]] <- as.matrix(Z_IC[[I]][[1]])
    z2_list[[i]] <- as.matrix(Z_IC[[I]][[2]])
    z3_list[[i]] <- as.matrix(Z_IC[[I]][[3]])
    i <- i + 1
}

Z_list <- c(list(z1_list),list(z2_list), list(z3_list))
# Monto Carlo estimators of Z
MZ123 <- NULL
for (l in 1:3) {
  MZ123[[l]] <- Reduce("+", Z_list[[l]]) / length(Z_list[[l]])
  MZ123[[l]] <- as.matrix(MZ123[[l]])
}
```

Fit the second clique: X45 | X3
```{r}
for(I in 1:M){
  SC[[I]] <- X_PIT[[I]][,c(3:5)]
  SC_fit[[I]] <- HTfit.model.output.dev(SC[[I]], U[[I]], init=rep(c(0.5,0.1),3))
  A_SC[[I]] <- SC_fit[[I]]$A
  B_SC[[I]] <- SC_fit[[I]]$B
  Z_SC[[I]] <- SC_fit[[I]]$Z
}

# obtain Monte Carlo estimators
MA_SC <- Reduce("+", A_SC) / length(A_SC)
MA_SC <- as.matrix(MA_SC)
MB_SC <- Reduce("+", B_SC) / length(B_SC)
MB_SC <- as.matrix(MB_SC)
```

```{r}
# obtain Monte Carlo estimators for Z
z33_list <- NULL
z4_list <- NULL
z5_list <- NULL
i <- 1
for (I in 1:M) {
    z33_list[[i]] <- as.matrix(Z_SC[[I]][[1]])
    z4_list[[i]] <- as.matrix(Z_SC[[I]][[2]])
    z5_list[[i]] <- as.matrix(Z_SC[[I]][[3]])
    i <- i + 1
}
Z2_list <- c(list(z33_list),list(z4_list),list(z5_list))
MZ345 <- NULL
for (l in 1:3) {
  MZ345[[l]] <- Reduce("+", Z2_list[[l]]) / length(Z2_list[[l]])
  MZ345[[l]] <- as.matrix(MZ345[[l]])
}
```

Construct the full beta matrix:
```{r}
B <- matrix(0, 5, 5)
B[1:3,1:3] <- MB_IC
B[3:5,3:5] <- MB_SC

B[4,1] <- B[3,1]
B[1,4] <- B[1,3]
B[5,1] <- B[3,1]
B[1,5] <- B[1,3]

B[4,2] <- B[3,2]
B[2,4] <- B[2,3]
B[5,2] <- B[3,2]
B[2,5] <- B[2,3]
B
```

Construct the full alpha matrix:
```{r}
A <- matrix(0, 5, 5)
A[1:3,1:3] <- MA_IC
A[3:5,3:5] <- MA_SC

A[4,1] <- A[4,3]*A[3,1]
A[1,4] <- A[1,3]*A[3,4]

A[5,1] <- A[5,3]*A[3,1]
A[1,5] <- A[1,3]*A[3,5]

A[4,2] <- A[4,3]*A[3,2]
A[2,4] <- A[2,3]*A[3,4]

A[5,2] <- A[5,3]*A[3,2]
A[2,5] <- A[2,3]*A[3,5]

A
```

For the case that simulation starts from C1:
```{r}
v <- 2
rQ1 <- function(N, v, A, B, Z){
    d <- ncol(A)
    X     <- matrix(nrow=N, ncol=d)
    Xstar <- NULL
    K      <- sample(1:d, size=N, replace=TRUE)
    Xstar <- lapply(as.list(K), function(k){
        alphak <- A[k,] 
        betak  <- B[k,]    
        Zk     <- Z[[k]]
        ## 
        stopifnot(all(Zk[,k]==0))
        h             <- NULL
        h             <- rep(0, d)
        h[-k]         <- apply(as.matrix(Zk[,-k]), 2, function(x) density(x, kernel="gaussian")$bw)
        Zkstar_edf    <- Zk[sample(1:nrow(Zk), 1, replace=TRUE), ]
        Zkstar_kernel <- apply(cbind(Zkstar_edf, h), 1, function(x) rnorm(1, mean=x[1], sd=x[2]))
        vplusE        <- v + rexp(1) 
        Xstar         <- ak(vplusE, alpha=alphak) + bk(vplusE, beta=betak)*Zkstar_kernel

        return(Xstar)
    })
    Xstar <- do.call("rbind", Xstar)
    
    # check for the simulated data in the separators and include the termination
    Xstar45 <- lapply(Xstar[,3], function(x) {
      if(x > v) { 
        alphak <- MA_IC[3,c(2:3)]
        betak <- MA_IC[3,c(2:3)]
        Z45 <- MZ345[[3]][,c(2:3)]
        h <- apply(Z45,2, function(x) density(x, kernel="gaussian")$bw)
        Zkstar_edf <- Z45[sample(1:nrow(Z45), 1, replace=TRUE),]
        Zkstar_kernel <- apply(cbind(Zkstar_edf, h), 1, function(x) rnorm(1, mean=x[1], sd=x[2]))
        Xstar45 <- ak(x, alpha=alphak) + bk(x, beta=betak)*Zkstar_kernel
        return(Xstar45)
      }  else{
        return(c(0, 0))  
        }
      })    
    Xstar45 <- do.call("rbind", Xstar45)
    
    X <- cbind(Xstar, Xstar45)

    out <- list()
    out$sim <- X
    out$ex <- K
    return(out)
}

SIM11 <- rQ1(10, v=v, A=MA_IC, B=MB_IC, Z=MZ123)
SIM111 <-as.data.frame(SIM11$sim) 
SIM11
```

For the case that simulation starts from C2:
```{r}
rQ2 <- function(N, v, A, B, Z){
    d <- ncol(A)
    X     <- matrix(nrow=N, ncol=d)
    Xstar <- NULL
    K      <- sample(1:d, size=N, replace=TRUE)
    Xstar <- lapply(as.list(K), function(k){
        alphak <- A[k,] 
        betak  <- B[k,]    
        Zk     <- Z[[k]]
 
        stopifnot(all(Zk[,k]==0))
        h             <- NULL
        h             <- rep(0, d)
        h[-k]         <- apply(as.matrix(Zk[,-k]), 2, function(x) density(x, kernel="gaussian")$bw)
        Zkstar_edf    <- Zk[sample(1:nrow(Zk), 1, replace=TRUE), ]
        Zkstar_kernel <- apply(cbind(Zkstar_edf, h), 1, function(x) rnorm(1, mean=x[1], sd=x[2]))
        vplusE        <- v + rexp(1) 
        Xstar         <- ak(vplusE, alpha=alphak) + bk(vplusE, beta=betak)*Zkstar_kernel

        return(Xstar)
    })
    Xstar <- do.call("rbind", Xstar)
    
    # check for the simulated data in the separators and include the termination
    Xstar12 <- lapply(Xstar[,1], function(x) {
      if(x > v) { 
        alphak <- MA_SC[3,c(1:2)]
        betak <- MA_SC[3,c(1:2)]
        Z12 <- MZ123[[3]][,c(1:2)]
        h <- apply(Z12,2, function(x) density(x, kernel="gaussian")$bw)
        Zkstar_edf <- Z12[sample(1:nrow(Z12), 1, replace=TRUE),]
        Zkstar_kernel <- apply(cbind(Zkstar_edf, h), 1, function(x) rnorm(1, mean=x[1], sd=x[2]))
        Xstar12 <- ak(x, alpha=alphak) + bk(x, beta=betak)*Zkstar_kernel
        return(Xstar12)
      }  else{
        return(c(0, 0))  
        }
      })    
    Xstar12 <- do.call("rbind", Xstar12)
    
    X <- cbind(Xstar12, Xstar)

    out <- list()
    out$sim <- X
    out$ex <- K
    return(out)
}

SIM2 <- rQ2(10, v=v, A=MA_SC, B=MB_SC, Z=MZ345)
SIM22 <-as.data.frame(SIM2$sim) 
SIM2
```













